Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2023-09-14 Paper/Manuscript.tex
Revised version of paper

\usepackage{lineno}
\linenumbers

@@

% All the \renewcommand stuff is now in a separate file so that it can be 
% reused for the list of symbols in the Supplementary document.
\input{symbol_aliases.tex}

@@

\author{Lars Nonboe Andersen}
\affiliation{Kongsberg Maritime AS, Strandpromenaden 50, 3191, Horten, Norway}

@@

\author{Harald Heimvoll}
\affiliation{Kongsberg Maritime AS, Strandpromenaden 50, 3191, Horten, Norway}

@@

\begin{abstract}
1. The use of quantitative broadband echosounders for biological studies and surveys offers considerable advantages over narrowband echosounders. These include improved spectral-based target identification and significantly increased ability to resolve individual targets. An understanding of current processing steps is required to fully utilize and further develop broadband acoustic methods in marine ecology.\linebreak 
2. We describe the steps involved in processing broadband acoustic data from raw data to frequency dependent target strength (TS(f)) and volume backscattering strength (Sv(f)) using data from the EK80 broadband scientific echosounder as examples. Although the overall processing steps are described and build on established methods from literature, multiple choices need to be made during implementation. \linebreak
3. To highlight and discuss some of these choices and facilitating a common understanding within the community, we have also developed a code which will be made publicly available and open source. The code follows the steps using raw data from two single pings, showing the step-by-step processing from raw data to TS(f) and Sv(f). \linebreak
4. This code can serve as a reference for developing own code or implementation in existing processing pipelines, as an educational tool and as a starting point for further development of broadband acoustic methods in fisheries acoustics.\linebreak

@@

Active acoustics is an efficient method for remote sensing of marine ecosystems and is used to cover a wide range of spatial and temporal scales, ranging from tens of km \citep{makris_fish_2006} to small scale behaviour patterns \citep{klevjer_split-beam_2003}, and can provide information on key forage species, including fish and zooplankton, linking primary producers and top predators as well as other taxonomic groups important for the ecosystem functioning \citep{benoit-bird_ecological_2016}. As early as 1935, \citet{sund_echo_1935} observed the distribution of spawning cod in the Lofoten area using a single beam echosounder. The method was further developed to map the abundance of fish, driven by the need for fisheries management \citep{Simmonds2005Fisheries}. More recently, active acoustics have been deployed on a wide range of platforms, including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.

Several different configurations of acoustic instruments are available, including multibeam sonars \citep{gerlotto_two_1999}, synthetic aperture sonar, omnidirectional sonars \citep{misund_improved_1996}, acoustic imaging sonars \citep{jaffe_ftv_1995} as well as single- and multibeam echosounders \citep{trenkel_new_2008}. Narrowband single beam echosounders have been used extensively in fisheries management and ecosystem monitoring, and the methodologies for using these systems are well developed. The ability for accurate calibration of the system has made these especially useful for providing data for fisheries management \citep{Simmonds2005Fisheries}.

Recent commercially available single beam echosounders produce pulses with a wide and continuous frequency range (broadband pulses), compared to conventional single frequency systems. This provides significantly better along-beam resolution, a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}, and improved frequency resolution for backscatter categorization \citep{korneliussen2018}.

Broadband echosounders have been actively investigated for ecological research, based on the expectation of improved categorization of species or group of species. Using high frequency broadband acoustics (150-600 kHz) \citet{lavery_measurements_2010} showed that broadband signals helped to reduce the ambiguities in interpretation of acoustic scattering from zooplankton and oceanic microstructure. \citet{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) that included the resonance frequency of swimbladdered fish, demonstrating both the spectral resolution and ability to classify size classes of fish within fish assemblages of broadband signals. \citet{blanluet_characterization_2019} used broadband acoustics, with ground truthing, to identify that the composition of two sound scattering layers (SSL) detected in the Bay of Biscay in springtime was able to be characterized, and also identified fine scale heterogeneity within the SSLs. While unsuccessful at discriminating between several species of fishes with large swimbladders during the Alaska pollock survey, \citet{bassett_broadband_2018} showed that broadband signals may be helpful in characterizing smaller fishes with swimbladders and euphausiids. \citet{benoit-bird_exploring_2020} was able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband signals (45-170 kHz), while attempts to classify species using multiple narrowband signals failed.

The increased spatial resolution is another aspect of broadband acoustics that is being actively used both for improved categorization and the increased ability to resolve targets near each other or near boundaries such as the seafloor. \citet{lavery2017} explored different broadband pulse shapes to increase the ability to resolve adjacent single targets as well as near boundaries in tank experiments. By applying high frequency broadband pulses to fish-like artificial targets, \citet{kubilius_remote_2020} demonstrated the potential for acoustic sizing of individually resolved fishes in a controlled ex situ environment. Utilizing the increased spatial resolution \citet{hasegawa_situ_2021} were able to isolate single fishes and discriminate successfully between average frequency responses of walleye pollock and pointhead flounder.

An equally interesting outcome from the increased resolution and broad frequency information is the ability to infer behavioural information for single individuals, groups of individuals, and interaction between individuals or groups. \citet{Traykovski1998Effect} was able to extract orientation information for krill by using a combination of high frequency broadband acoustics and acoustic modelling. Fine-scale observations of single targets facilitated the study of interactions between prey and predators (capelin and cod) in the Barents sea \citep{skaret_diel_2020}.

There have been several scientific broadband echosounder systems developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}. Commercially available systems are now produced by several manufacturers.

The theoretical foundation for signal processing comes from radar applications and has been documented for broadband echosounders \citep{stanton2008}. To tap the full potential of broadband echosounder data, further data processing methods are being developed \citep{lavery2017,bassett_broadband_2018}. Examples include adapting pulses to improve target separation and obtain additional parameters to describe the individual targets. Different transmission pulse forms can be envisioned as well as various methods for acoustic target classification. When translating equations into computer processing code, several choices must be made - some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature. The latter is typically missing in the literature, making it difficult to test the implementation of the signal processing methods in new echo sounders and post processing software. By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in marine ecosystem acoustics.

The objective of this paper and associated code is to present a systematic and comprehensive description of the data processing steps for calibrated broadband echosounder data. The steps include pulse compression, target strength as a function of frequency [$\ts(\freqsym)$], and volume backscattering strength as a function of frequency [$\sv(\freqsym)$]. The intention is that the code will be used as a starting point for implementations in various relevant data processing software and for further developing active acoustic broadband signal processing, and to serve as a learning resource.

@@

\subsection{Accompanying code}
The code accompanying this paper is written in the Python programming language and is available in the supplementary materials. All single ping processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}). Reproduction of figures 9 and 12 requires the original echosounder raw data and are not included. Without loss of generality, we use the \ek{} echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.

Our presentation uses nomenclature and approaches that are commonly used for narrowband echosounder systems, which were derived from radar processing \citep{cook1967}. In particular, the expressions for target strength ($\ts$) and volume backscattering strength ($\sv$) \citep{MacLennan2002consistent} are presented in a similar manner for broadband signals as for narrowband signals.

\subsection{System overview}
A basic quantitative echosounder system consists of a transducer, a transceiver, and a computer program that controls the operation of the transceiver and records received signals. During transmission the program defines the signals that are created as electric signals in the transceiver, converted to acoustic signals by the transducer and transmitted into the water. The acoustic signals propagate through the water, are reflected or scattered by objects in the water, and propagate back to the transducer. During reception the transducer converts the received acoustic signals to electric signals, which are received, pre-amplified, filtered, digitized, processed in the transceiver, and then transferred to the controlling program for further data processing and storage (Fig. \ref{fi:ek_sys}). Many types of transmit signals are feasible - this paper considers only linear frequency modulated signals (also known as linear chirps).

\begin{figure}
\includegraphics[width=16cm]{Fig_ek_sys}
\caption{\label{fi:ek_sys}Signal and data flow in the \ek{} system. An echosounder ping starts with the definition of a transmit signal (upper left) and ends with file storage (lower middle) and display and analysis (lower right).}
\end{figure}

@@

The controlling computer program generates a short-duration digital transmit signal (a ping), $\ytx(\samplesymt)$, where $\samplesymt$ is the sample index in the discrete time domain. Typical broadband pulses are linear upsweep pulses windowed by an envelope function. The generated signal is converted to an analogue electric signal $\ytxe(\timesym)$ and amplified by the transceiver to obtain the analogue signal $\ytxe(\timesym)$, where $\timesym$ is the time for the signal. The analogue and amplified signals are passed on to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ in the water. For a splitbeam echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.

In the example a linear sweep enveloped by a Hanning window is implemented (Fig.~\ref{fi:ytx}), where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively. A slope of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
\begin{figure}
\includegraphics[width=16cm]{Fig_ytx}
\caption{\label{fi:ytx} Envelope of linear chirp pulses from 92 kHz to 158 kHz with a slope of 0.057 (blue) and 0.5 (orange).}
\end{figure}

@@

To remove noise and reduce the quantity of data, the sampled signal from each channel is filtered and decimated in multiple stages, $\stagesym$, using complex bandpass filters, $\hfl(\genidxsym,\stagesym)$, and decimation factors, $\decfac(\stagesym)$. The individual filter coefficients for each filter and decimation stage are indexed by $\genidxsym$. The output signal from each channel, $\channelsym$, from each filter and decimation stage, $\stagesym$, is then given by:
%
\begin{equation}
\label{eq:yrx}
\yrx(\samplesymt,\channelsym,\stagesym) = \left( \yrx(\samplesymt,\channelsym,\stagesym-1) * \hfl(\genidxsym,\stagesym) \right)_{\downarrow \decfac(\stagesym)}, 
\stagesym = 1,\ldots,\nstages,
\end{equation}
%
where $\yrx(\samplesymt,\channelsym,0)$ is set to $\yrxorg(\samplesymt,\channelsym)$, being the signal before decimation, $*$ indicates convolution, $\downarrow$ indicates decimation by the factor $\decfac(\stagesym)$, and $\nstages$ is the total number of filter stages. The output signal from the final filter and decimation stage, $\yrx(\samplesymt,\channelsym,\nstages)$, is shortened to $\yrx(\samplesymt,\channelsym)$ for convenience. For the output signal, $\yrx(\samplesymt,\channelsym)$, the decimated sampling rate, $\fsdec$, is given by:
%
\begin{equation}
\label{eq:fsdec}
\fsdec = \fs\prod_{\stagesym=1}^{\nstages} \frac{1}{\decfac(\stagesym)}.
\end{equation}
%
The characteristics of the bandpass filter and decimation factors are chosen with regard to the desired operating bandwidth, noise suppression levels, impulse response duration, and other common filter characteristics, with the aim of maintaining sufficient information in the data. In the example $N_v=2$. The frequency responses of the filters are shown in Figure~\ref{fi:fir} and the corresponding filter coefficients and decimation factors are given in the test data set.

\begin{figure}
\includegraphics[width=16cm]{Fig_fir}
\caption{\label{fi:fir} The frequency response (filter gain) of the filters in our test set. The blue and orange curve represents the filter response of the first and second filter. The horizontal green line indicates the frequency range of the transmit signal.}
\end{figure}

@@

\subsection{Pulse compression}
To increase signal-to-noise ratio and resolution along the acoustic beam a matched filter may be applied to the raw data samples \citep{turin1960}. This technique is also known as pulse compression \citep{klauder1960}. One approach for a matched filter is to use a normalized version of the ideal transmit signal as the replicate signal, filtered and decimated using the same filters and decimation factors as applied in Eq. \ref{eq:yrx}. The normalized ideal transmit signal, $\ytxnorm(\samplesymt)$, is given by:
%
\begin{equation}
\label{eq:ytxnorm}
\ytxnorm(\samplesymt) = \frac{\ytx(\samplesymt)}{\textrm{max}(\ytx(\samplesymt))}\end{equation}
%
where $\textrm{max}$ is the maximum value of $\ytx(\samplesymt)$. The filtered and decimated output signal, $\ytxnorm(\samplesymt,\stagesym)$, from each filter stage, $\stagesym$, using the normalized ideal transmit signal, $\ytxnorm(\samplesymt)$, as the input signal, is given by:
%
\begin{equation}
\label{eq:FilterStagesTX}
\ytxnorm(\samplesymt,\stagesym) = \left[ \ytxnorm(\samplesymt,\stagesym-1) * \hfl(\genidxsym,\stagesym) \right]_{\downarrow \decfac(\stagesym)}, 
\stagesym = 1,\ldots,\nstages,
\end{equation}
%
where $\ytxnorm(\samplesymt,0)$ is set to $\ytxnorm(\samplesymt)$. The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$  (Fig.~\ref{fi:y_mf_n}).
%
\begin{figure}
\includegraphics[width=16cm]{Fig_y_mf_n}
\caption{\label{fi:y_mf_n} The absolute value of the filtered and decimated output signal, $\ymf(\samplesymt)$, that is used for the pulse compression.}
\end{figure}

The auto correlation function of the matched filter signal and the effective pulse duration, defined as the pulse duration at transmit power $\ptxe$ which produces the same energy as the actual transmitted pulse, will be used in later processing steps and are defined as
\begin{equation}
\label{eq:TXAuto}
\ymfauto(\samplesymt) = \frac{\ymf(\samplesymt)*\ymf^*(-\samplesymt)}{||\ymf||^2_2}
\end{equation}
and
\begin{equation}
\label{eq:TauEff}
\teff = \frac{\sum \ptxauto(\samplesymt)}{\textrm{max}(\ptxauto(\samplesymt))\fsdec},
\end{equation}
where
\begin{equation*}
%\label{eq:TauEff}
\ptxauto(\samplesymt)  =  |\ymfauto(\samplesymt)|^2
\end{equation*}
is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over a duration of twice the nominal pulse duration, $2\tnom$.

\begin{figure}
\includegraphics{Fig_ACF}
\caption{The autocorrelation function $\ptxauto$ from the example. The corresponding $\teff=0.1$ ms and $\tnom=2$ ms\label{fi:ACF}.}
\end{figure}

>>>>

\usepackage{lineno}
\usepackage{xurl}

\linenumbers

@@

\newcommand{\ek}{Simrad EK80}

\newcommand{\timesym}{t}
\newcommand{\freqsym}{f}
\newcommand{\samplesymt}{n}
\newcommand{\samplesymf}{m}
\newcommand{\genidxsym}{i}

\newcommand{\channelsym}{u}
\newcommand{\nchannels}{N_{\textrm{u}}}

\newcommand{\stagesym}{v}
\newcommand{\nstages}{N_{\textrm{v}}}

\newcommand{\fs}{f_{\textrm{s}}}
\newcommand{\fsdec}{f_{\textrm{s,dec}}}
\newcommand{\fstart}{f_{\textrm{start}}}
\newcommand{\fstop}{f_{\textrm{stop}}}
\newcommand{\fc}{f_{\textrm{c}}}
\newcommand{\fn}{f_{\textrm{n}}}

\newcommand{\zrxe}{z_{\textrm{rx,e}}}
\newcommand{\ztde}{z_{\textrm{td,e}}}

\newcommand{\ptxe}{p_{\textrm{tx,e}}}
\newcommand{\prxe}{p_{\textrm{rx,e}}}

\newcommand{\ntd}{n_{\textrm{td}}}
\newcommand{\tnom}{\tau}
\newcommand{\teff}{\tau_{\textrm{eff}}}

\newcommand{\ytxe}{y_{\textrm{tx,e}}}
\newcommand{\ytxa}{y_{\textrm{tx,a}}}
\newcommand{\yrxa}{y_{\textrm{rx,a}}}
\newcommand{\yrxe}{y_{\textrm{rx,e}}}

\newcommand{\ytx}{y_{\textrm{tx}}}
\newcommand{\ytxnorm}{\tilde{y}_{\textrm{tx}}}
\newcommand{\ytxfd}{y_{\textrm{tx,f,d}}}
\newcommand{\yrx}{y_{\textrm{rx}}}
\newcommand{\yrxorg}{y_{\textrm{rx,org}}}
\newcommand{\ymf}{y_{\textrm{mf}}}

\newcommand{\ytd}{y_{\textrm{td}}}

\newcommand{\ypc}{y_{\textrm{pc}}}
\newcommand{\ypctarget}{y_{\textrm{pc,t}}}
\newcommand{\ypcspread}{y_{\textrm{pc,s}}}
\newcommand{\ymfauto}{y_{\textrm{mf,auto}}}
\newcommand{\ymfautored}{y_{\textrm{mf,auto,red}}}

\newcommand{\ptxauto}{p_{\textrm{tx,auto}}}

\newcommand{\ypctargetf}{Y_{\textrm{pc,t}}}
\newcommand{\ypctargetnormf}{\tilde{Y}_{\textrm{pc,t}}}
\newcommand{\ypcvolumef}{Y_{\textrm{pc,v}}}
\newcommand{\ypcvolumenormf}{\tilde{Y}_{\textrm{pc,v}}}

\newcommand{\ymfautof}{Y_{\textrm{mf,auto}}}
\newcommand{\ymfautoredf}{Y_{\textrm{mf,auto,red}}}

\newcommand{\prxetf}{P_{\textrm{rx,e,t}}}
\newcommand{\prxevf}{P_{\textrm{rx,e,v}}}

\newcommand{\fstage}{k}
\newcommand{\decfac}{D}
\newcommand{\lfl}{L_{\textrm{fl}}}
\newcommand{\hfl}{h_{\textrm{fl}}}
\newcommand{\hbp}{h_{\textrm{bp}}}
\newcommand{\hannw}{w}
\newcommand{\hannwnorm}{\tilde{\hannw}}
\newcommand{\nw}{N_{\hannw}}
\newcommand{\hannwpart}{\gamma}
\newcommand{\tslide}{t_w}

\newcommand{\bs}{\sigma_{\textrm{bs}}}
\newcommand{\mysp}{S_p}
\newcommand{\ts}{\textrm{TS}}
\newcommand{\sv}{S_{\textrm{v}}}

\newcommand{\range}{r}
\newcommand{\rangeref}{r_0}
\newcommand{\athw}{\phi}
\newcommand{\along}{\theta}
\newcommand{\gain}{g}
\newcommand{\gainzero}{g_0}
\newcommand{\eqang}{\psi}

\newcommand{\rtarget}{r_{\textrm{target}}}
\newcommand{\alongtarget}{\theta_{\textrm{target}}}
\newcommand{\athwtarget}{\phi_{\textrm{target}}}

\newcommand{\wlen}{\lambda}
\newcommand{\cw}{c}
\newcommand{\absorp}{\alpha}

\newcommand{\dft}{\textrm{DFT}}
\newcommand{\ndft}{{N_{\textrm{DFT}}}}
\newcommand{\ndftw}{_{\nw}}
\newcommand{\atan}{\textrm{arctan2}}
\newcommand{\anglefalong}{\gamma_\along}
\newcommand{\anglefathw}{\gamma_\athw}

\newcommand{\sigmabs}{\sigma_{\textrm{bs}}}

\newcommand{\code}[1]{\texttt{#1}}

@@

\author{Lars Nonboe Andersen}
\affiliation{Kongsberg Discovery AS, Strandpromenaden 50, 3191, Horten, Norway}

@@

\author{Harald Heimvoll}
\affiliation{Kongsberg Discovery AS, Strandpromenaden 50, 3191, Horten, Norway}

@@

\begin{abstract}
1. The use of quantitative broadband echosounders for biological studies and surveys can offer considerable advantages over narrowband echosounders. These include improved spectral-based target identification and significantly increased ability to resolve individual targets. An understanding of current processing steps is required to fully utilize and further develop broadband acoustic methods in marine ecology.\linebreak 
2. We describe the steps involved in processing broadband acoustic data from raw data to frequency dependent target strength (TS(f)) and volume backscattering strength (Sv(f)) using data from the EK80 broadband scientific echosounder as examples. Although the overall processing steps are described and build on established methods from the literature, multiple choices need to be made during implementation. \linebreak
3. To highlight and discuss some of these choices and facilitate a common understanding within the community, we have also developed a Python code which will be made publicly available and open source. The code follows the steps using raw data from two single pings, showing the step-by-step processing from raw data to TS(f) and Sv(f). \linebreak
4. This code can serve as a reference for developing custom code or implementation in existing processing pipelines, as an educational tool and as a starting point for further development of broadband acoustic methods in fisheries acoustics.\linebreak

@@

Echosounders are used for remote sensing marine ecosystems. As early as 1935, \citet{sund_echo_1935} observed the distribution of spawning cod in the Lofoten area using a single beam echosounder. The method was further developed to map the abundance of fish, driven by the need for fisheries management \citep{Simmonds2005Fisheries}. More recently, fisheries acoustics sensors have been deployed on a wide range of platforms, including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.

Today echosounders can produce pulses with a wide and continuous frequency range (broadband pulses), compared to the conventional narrowband systems. This provides in most cases significantly better along-beam resolution, a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}, and improved frequency resolution for backscatter categorization \citep{korneliussen2018}.  \citet{lavery_measurements_2010} used broadband signals to reduce the ambiguities in interpretation of acoustic scattering from zooplankton and oceanic microstructure. \citet{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) that included the resonance frequency of swimbladdered fish to classify size classes. \citet{blanluet_characterization_2019} used broadband acoustics, with ground truthing, to characterize the composition of two sound scattering layers (SSL). \citet{bassett_broadband_2018} showed that broadband signals may be helpful in characterizing smaller fishes with swimbladders and euphausiids. \citet{benoit-bird_exploring_2020} were able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband signals (45-170 kHz). \citet{lavery2017} explored different broadband pulse shapes to increase the ability to resolve adjacent single targets as well as near boundaries in tank experiments. By applying high frequency broadband pulses to fish-like artificial targets, \citet{kubilius_remote_2020} demonstrated the potential for acoustic sizing of individually resolved fishes in a controlled ex situ environment. Utilizing the increased spatial resolution \citet{hasegawa_situ_2021} were able to isolate single fishes and discriminate successfully between average frequency responses of walleye pollock and pointhead flounder.

Several scientific broadband echosounder systems have been developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}. More widespread use of broadband acoustics in fisheries and ecosystem research have followed from the upgrade of widely used narrowband systems to gain broadband capabilities.

Signal processing methods for broadband echosounders \citep{stanton2008} are based on radar signal processing theory, with further adaptation to echosounders \citep{lavery2017,bassett_broadband_2018}. When implementing the equations to computer processing code, some choices are well founded in the signal processing literature whereas others are of a more practical and ad-hoc nature. The latter is typically missing in the literature, making it difficult to benchmark new methods and to test the implementation in new echosounders and post processing software.

The objective of this paper and associated code is to provide a benchmark for developing and implementing signal processing algorithms for broadband echosounder. We present the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data. The steps include pulse compression, target strength as a function of frequency ($\ts(\freqsym)$, dB re 1m$^2$), and volume backscattering strength as a function of frequency ($\sv(\freqsym)$, dB re 1m$^-1$). The intention is that the code will be used as a starting point for implementations in various relevant data processing software, for further developing active acoustic broadband signal processing, and to serve as a learning resource.

@@

\subsection{Accompanying code}
The code  accompanying this paper is written in the Python programming language (v3.10) and is available through GitHub (see \ref{dataavstatement} for code and data availability) together with the data used in the examples. All single ping processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}). Reproduction of Fig. \ref{fi:ts_theta_phi}a and Fig. \ref{fi:Fig_Sv_echogram}a requires downloading the original echosounder raw data separately and using the scripts (\code{TSfEchogram.py} and \code{SvfEchogram.py}). Without loss of generality, we use the \ek{} echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field. The test data sets accompanying the code are representative for the data contained in Simrad EK80 raw files. In order to make the code echosounder independent simple json strings are used as input. Detailed information on settings used during data collection are available in the code and the data files on GitHub.

Our presentation uses nomenclature and approaches that are commonly used for narrowband echosounder systems, which were derived from radar processing \citep{cook1967}. In particular, the expressions for target strength ($\ts$, dB re 1m${^2}$) and volume backscattering strength ($\sv$, dB re 1m$^{-1}$) \citep{MacLennan2002consistent} are presented in a similar manner for broadband signals as for narrowband signals.

\subsection{System overview}
A basic quantitative echosounder system consists of a transducer, a transceiver, and a computer program that controls the operation of the transceiver and records received signals. During transmission the program defines the signals that are created as electric signals in the transceiver, converted to acoustic signals by the transducer and transmitted into the water. The acoustic signals propagate through the water, are reflected or scattered by objects in the water, and propagate back to the transducer. During reception the transducer converts the received acoustic signals to electric signals, which are received, pre-amplified, filtered, digitized, processed in the transceiver, and then transferred to the controlling program for further data processing and storage (Fig. \ref{fi:ek_sys}). Many types of transmit signals are feasible - this paper considers only upsweep linear frequency modulated signals (also known as linear chirps).

\begin{figure}
\includegraphics[width=16cm]{Fig_ek_sys}
\caption{\label{fi:ek_sys} Signal and data flow in the \ek{} system. An echosounder ping starts with the definition of a transmit signal (upper left) and ends with file storage and display and analysis after matched filtering.}
\end{figure}

@@

The controlling computer program generates a short-duration digital transmit signal (a ping), $\ytx(\samplesymt)$, where $\samplesymt$ is the sample index in the discrete time domain. The nominal pulse duration, $\tnom$, is defined as the duration of the digital transmit signal $\ytx(\samplesymt)$.
Typical broadband pulses are linear upsweep pulses windowed by an envelope function. The generated signal is converted to an analogue electric signal $\ytxe(\timesym)$ and amplified by the transceiver to obtain the analogue signal $\ytxe(\timesym)$, where $\timesym$ is the time for the signal. The analogue and amplified signals are passed on to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ in the water. For a splitbeam echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.

The most commonly used transmit signals (e.g. in Simrad EK80) are linear frequency modulated signals with an applied Tukey window function (see e.g. \citet{lavery2017} for a more in-depth treatment of this subject).

@@

To remove noise and reduce the quantity of data, the sampled signal from each channel is filtered and decimated in multiple stages, $\stagesym$, using complex bandpass filters, $\hbp(\genidxsym,\stagesym)$, and decimation factors, $\decfac(\stagesym)$. This processing step is approximately equivalent to a traditional demodulation process. See e.g. Table 1.2 in \citep{Demer2017} for examples of values. The individual filter coefficients for each filter and decimation stage are indexed by $\genidxsym$. The output signal from each channel, $\channelsym$, from each filter and decimation stage, $\stagesym$, is then given by:
%
\begin{equation}
\label{eq:yrx}
\yrx(\samplesymt,\channelsym,\stagesym) = \left( \yrx(\samplesymt,\channelsym,\stagesym-1) * \hbp(\genidxsym,\stagesym) \right)_{\downarrow \decfac(\stagesym)}, 
\stagesym = 1,\ldots,\nstages,
\end{equation}
%
where $\yrx(\samplesymt,\channelsym,0)$ is set to $\yrxorg(\samplesymt,\channelsym)$, being the signal before decimation, $*$ indicates convolution, $\downarrow$ indicates decimation by the factor $\decfac(\stagesym)$, and $\nstages$ is the total number of filter stages. The output signal from the final filter and decimation stage, $\yrx(\samplesymt,\channelsym,\nstages)$, is shortened to $\yrx(\samplesymt,\channelsym)$ for convenience. For the output signal, $\yrx(\samplesymt,\channelsym)$, the decimated sampling rate, $\fsdec$, is given by:
%
\begin{equation}
\label{eq:fsdec}
\fsdec = \fs\prod_{\stagesym=1}^{\nstages} \frac{1}{\decfac(\stagesym)}.
\end{equation}
%
The characteristics of the bandpass filter and decimation factors are chosen with regard to the desired operating bandwidth, noise suppression levels, impulse response duration, and other common filter characteristics, with the aim of maintaining sufficient information in the data \citep{ProakisDSP,CrochireMDSP}. The frequency responses of the filters are shown in Fig. ~\ref{fi:fir} and the corresponding filter coefficients and decimation factors are given in the test data set, where $N_v=2$.

\begin{figure}
\includegraphics[width=16cm]{Fig_fir}
\caption{\label{fi:fir} Example of frequency response (filter gain) of the filters in our test set. The blue and orange curve represents the filter response of the first and second filter. Note that the blue line is above 0 dB, this is caused by the transition from complex to real values. The vertical dashed green lines indicates the frequency range of the transmit signal.}
\end{figure}

@@

\subsection{Matched filtering (pulse compression)}
To increase signal-to-noise ratio and resolution along the acoustic beam a matched filter may be applied to the raw data samples \citep{turin1960}. This technique is also known as pulse compression \citep{klauder1960}, achieving the average transmitted power of a relatively long pulse while obtaining the range resolution of a shorter pulse. One approach for a matched filter is to use a normalized version of the ideal transmit signal as the replicate signal, filtered and decimated using the same filters and decimation factors as applied in Eq. \ref{eq:yrx}. The normalized ideal transmit signal, $\ytxnorm(\samplesymt)$, is given by:
%
\begin{equation}
\label{eq:ytxnorm}
\ytxnorm(\samplesymt) = \frac{\ytx(\samplesymt)}{\textrm{max}(\ytx(\samplesymt))}\end{equation}
%
where $\textrm{max}$ is the maximum value of $\ytx(\samplesymt)$. The filtered and decimated output signal, $\ytxnorm(\samplesymt,\stagesym)$, from each filter stage, $\stagesym$, using the normalized ideal transmit signal, $\ytxnorm(\samplesymt)$, as the input signal, is given by:
%
\begin{equation}
\label{eq:FilterStagesTX}
\ytxnorm(\samplesymt,\stagesym) = \left[ \ytxnorm(\samplesymt,\stagesym-1) * \hbp(\genidxsym,\stagesym) \right]_{\downarrow \decfac(\stagesym)}, 
\stagesym = 1,\ldots,\nstages,
\end{equation}
%
where $\ytxnorm(\samplesymt,0)$ is set to $\ytxnorm(\samplesymt)$. The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$  (Fig.~\ref{fi:y_mf_n_ACF}a).
%
\begin{figure}
\includegraphics[width=16cm]{Fig_y_mf_n_ACF}
\caption{\label{fi:y_mf_n_ACF} An example where the decimated sampling rate ($\fsdec$, Eq. \ref{eq:fsdec}) is 125 kHz, the nominal pulse duration ($\tnom$) is 2 ms, and the effective pulse duration ($\teff$, Eq. \ref{eq:TauEff}) is 0.01 ms. (a) The absolute value of the filtered and decimated output signal, $\ymf(\samplesymt)$, from the final filter and decimation stage, that is used for the pulse compression. (b) The autocorrelation function $\ptxauto(n)$.}
\end{figure}

The auto correlation function of the matched filter signal ($\ymfauto(\samplesymt)$) and the effective pulse duration ($\teff$) will be used in later processing steps.

$\ymfauto(\samplesymt)$ is defined as

\begin{equation}
\label{eq:TXAuto}
\ymfauto(\samplesymt) = \frac{\ymf(\samplesymt)*\ymf^*(-\samplesymt)}{||\ymf||^2_2}
\end{equation}

where $"*"$ denotes convolution, $"^*"$ complex conjugate, and $||\ymf||_2$ the $l^2$-norm of $\ymf$, also known as the Euclidean norm (see e.g. \citet{PadgettSigProc} and \citet{GhatakR} for background on $l^2$-norm).

$\teff$ is defined as

\begin{equation}
\label{eq:TauEff}
\teff = \frac{\sum \ptxauto(\samplesymt)}{\textrm{max}(\ptxauto(\samplesymt))\fsdec},
\end{equation}

where

\begin{equation*}
%\label{eq:TauEff}
\ptxauto(\samplesymt)  =  |\ymfauto(\samplesymt)|^2
\end{equation*}

is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over the duration of the autocorrelation function (Fig. \ref{fi:y_mf_n_ACF}b).

To perform pulse compression the received signal, $\yrx(\samplesymt,\channelsym)$, is convolved with a complex conjugated and time-reversed version of the matched filter signal, and here also normalized with the $l^2$-norm of the matched filter to maintain received signal power. The pulse compressed signal, $\ypc(\samplesymt,\channelsym)$, then becomes
\begin{equation}
\label{eq:PulseComp}
\ypc(\samplesymt,\channelsym) = \frac{\yrx(\samplesymt,\channelsym)*\ymf^*(-\samplesymt)}{||\ymf||^2_2},
\end{equation}
%
The received power samples are then used to estimate target strength and volume backscattering strength. For estimating received power samples, the average signal, $\ypc(\samplesymt)$, over all transducer sectors, $\nchannels$, is used:

\begin{equation}
\label{eq:SumSig}
\ypc(\samplesymt) = \frac{1}{\nchannels} \sum_{\channelsym = 1}^{\nchannels} \ypc(\samplesymt,\channelsym).
\end{equation}
